---
title: 'KDE_HexBin_EstKFxn'
author: "Charysse Gibson"
date: "February 26, 2020"
output: pdf_document
---

```{r setup, include=F}
library(spatstat)
library(sf)
library(sp)
library(tmaptools)
library(tmap)
library(fMultivar)
library(maptools)
library(MASS)
```

# bramblecanes data

**1. Consider the bramblecanes data from class. For each of the three ages of bramble canes (0, 1, or 2) construct the following:**


a) a kernel density estimate and display it with a heat map (use trial and error to get an appropriate bin width)
b) a hexagonal bin plot
c) an estimate of the K function along with the envelope for testing CSR.


### kde heat maps
```{r bramblecanes kde heat maps, echo=TRUE, results=FALSE}
## Kernel Density Estimate Heat Maps
data(bramblecanes)
str(bramblecanes)
marks = bramblecanes$marks
x0 <- bramblecanes$x[bramblecanes$marks==0]
y0 <- bramblecanes$y[bramblecanes$marks==0]
x1 <- bramblecanes$x[bramblecanes$marks==1]
y1 <- bramblecanes$y[bramblecanes$marks==1]
x2 <- bramblecanes$x[bramblecanes$marks==2]
y2 <- bramblecanes$y[bramblecanes$marks==2]
```
#### Bramble canes 0 yrs
```{r bramblecanes kde heat maps 0 yrs}
bc0.kde <- kde2d(x0, y0, 0.3)
filled.contour(bc0.kde, xlim=c(0,1), ylim=c(0,1), asp=1)
```
#### Bramble canes 1 yrs
```{r bramblecanes kde heat maps 1 yrs}
bc1.kde <- kde2d(x1, y1, 0.3)
filled.contour(bc1.kde, xlim=c(0,1), ylim=c(0,1), asp=1 )
```
#### Bramble canes 2 yrs
```{r bramblecanes kde heat maps 2 yrs}
bc2.kde <- kde2d(x2, y2, 0.3)
filled.contour(bc2.kde, xlim=c(0,1), ylim=c(0,1), asp=1)
```


### hexagonal binning - 0 years

```{r bramblecanes hex bin yr0, echo=TRUE, results=FALSE}
## Hexagonal binning

yr0 <- bramblecanes[bramblecanes$marks==0]
yr0_sp <- as(yr0, Class = 'SpatialPoints')

spdf = yr0_sp
coordinates(spdf)
hbins = fMultivar::hexBinning(coordinates(spdf))

hexbin_map = function( spdf , ... )
{
  hbins = fMultivar::hexBinning( coordinates(spdf) , ... )
  u = c( 1 , 0 , -1 , -1 , 0 , 1 )
  v = c( 1 , 2 , 1 , -1 , -2 , -1 )
  u = u * min( diff( unique( sort( hbins$x))))
  v = v * min( diff( unique( sort( hbins$y))))/3
  hexes_list = vector( length(hbins$x) , mode="list" )
  for ( i in 1:length(hbins$x) )
  {
    pol= Polygon( cbind( u + hbins$x[i] , v + hbins$y[i] ) , 
                  hole=FALSE )
    hexes_list[[i]] = Polygons( list( pol) , i )
  }
  ## Note error in book for next line (correcet on web site)
  hex_cover_sp <- SpatialPolygons( hexes_list, 
                                   proj4string=CRS(proj4string(spdf)))
  hex_cover <- SpatialPolygonsDataFrame( hex_cover_sp ,
                       data.frame( z=hbins$z , match.ID=FALSE ) )
  return( hex_cover )
}

bc_hex <- hexbin_map(yr0_sp,bins=20)
bc0 <- 
  tm_shape(bc_hex) +
  tm_layout(main.title='bramble canes 0 yrs') +
  tm_fill(col="z", legend.z=0, title="Count", style='fixed', breaks=1:8, alpha=0.7)
```

### hexagonal binning - 1 years

```{r bramblecanes hex bin yr1, echo=TRUE, results=FALSE}
## Hexagonal binning

yr1 <- bramblecanes[bramblecanes$marks==1]
yr1_sp <- as(yr1, Class = 'SpatialPoints')

spdf = yr1_sp
coordinates(spdf)
hbins = fMultivar::hexBinning(coordinates(spdf))

hexbin_map = function( spdf , ... )
{
  hbins = fMultivar::hexBinning( coordinates(spdf) , ... )
  u = c( 1 , 0 , -1 , -1 , 0 , 1 )
  v = c( 1 , 2 , 1 , -1 , -2 , -1 )
  u = u * min( diff( unique( sort( hbins$x))))
  v = v * min( diff( unique( sort( hbins$y))))/3
  hexes_list = vector( length(hbins$x) , mode="list" )
  for ( i in 1:length(hbins$x) )
  {
    pol= Polygon( cbind( u + hbins$x[i] , v + hbins$y[i] ) , 
                  hole=FALSE )
    hexes_list[[i]] = Polygons( list( pol) , i )
  }
  ## Note error in book for next line (correcet on web site)
  hex_cover_sp <- SpatialPolygons( hexes_list, 
                                   proj4string=CRS(proj4string(spdf)))
  hex_cover <- SpatialPolygonsDataFrame( hex_cover_sp ,
                       data.frame( z=hbins$z , match.ID=FALSE ) )
  return( hex_cover )
}

bc_hex <- hexbin_map(yr1_sp,bins=20)
bc1 <- 
  tm_shape(bc_hex) +
  tm_layout(main.title='bramble canes 1 yrs') +
  tm_fill(col="z", legend.z=0, title="Count", style='fixed', breaks=1:8, alpha=0.7)
```

### hexagonal binning - 2 years

```{r bramblecanes hex bin yr2, echo=TRUE, results=FALSE}
## Hexagonal binning

yr2 <- bramblecanes[bramblecanes$marks==2]
yr2_sp <- as(yr2, Class = 'SpatialPoints')

spdf = yr2_sp
coordinates(spdf)
hbins = fMultivar::hexBinning(coordinates(spdf))

hexbin_map = function( spdf , ... )
{
  hbins = fMultivar::hexBinning( coordinates(spdf) , ... )
  u = c( 1 , 0 , -1 , -1 , 0 , 1 )
  v = c( 1 , 2 , 1 , -1 , -2 , -1 )
  u = u * min( diff( unique( sort( hbins$x))))
  v = v * min( diff( unique( sort( hbins$y))))/3
  hexes_list = vector( length(hbins$x) , mode="list" )
  for ( i in 1:length(hbins$x) )
  {
    pol= Polygon( cbind( u + hbins$x[i] , v + hbins$y[i] ) , 
                  hole=FALSE )
    hexes_list[[i]] = Polygons( list( pol) , i )
  }
  ## Note error in book for next line (correcet on web site)
  hex_cover_sp <- SpatialPolygons( hexes_list, 
                                   proj4string=CRS(proj4string(spdf)))
  hex_cover <- SpatialPolygonsDataFrame( hex_cover_sp ,
                       data.frame( z=hbins$z , match.ID=FALSE ) )
  return( hex_cover )
}

bc_hex <- hexbin_map(yr2_sp,bins=20)
bc2 <- 
  tm_shape(bc_hex) +
  tm_layout(main.title='bramble canes 2 yrs') +
  tm_fill(col="z", legend.z=0, title="Count", style='fixed', breaks=1:8, alpha=0.7)
```

```{r bramblecanes hex bins, warning=FALSE}
tmap_mode("plot")
tmap_arrange(bc0,bc1,bc2)
```

### k function estimate and envelope for testing CSR

```{R kf env CSR testing, echo=TRUE, results=FALSE}
kf.env.yr0 <- envelope(yr0, Kest, correction="border")
kf.env.yr1 <- envelope(yr1, Kest, correction="border")
kf.env.yr2 <- envelope(yr2, Kest, correction="border")

plot(kf.env.yr0)
plot(kf.env.yr1)
plot(kf.env.yr2)
```

# trees

**2. Consider the location of trees in the trees.csv data set given on Blackboard.**


a) Create a ppp object that has region [0,200] by [0,200].
b) The tree diameters are given in the variable dia. Make a histogram for the tree diameters.
c) Construct a kernel density and display it with a heat map.
d) ~~Display the kernel density with a set of contours.~~
e) Estimate the K function and plot it along with the envelope for testing CSR.
f) Consider only the trees with diameter less than or equal to 20. Repeat part (e).

```{r trees,  echo=TRUE, results=FALSE, warning=FALSE}
trees <- 
  read.csv("C:/Users/chary/OneDrive/Documents/SLU/Sping 2020/BST 6200 Spatial Statistics and Disease Mapping/Homework/HW2_data/trees.csv")

trees_ppp <- ppp(trees$x, trees$y, c(0,200), c(0,200))

hist(trees$dia)

trees_sp <- as.SpatialPoints.ppp(trees_ppp)
my_bw = 1 * c(1,1)
trees_dens = smooth_map(trees_sp, unit='feet', unit.size=12,
                        bandwidth = my_bw)
tm_shape(trees_dens$raster) + 
  tm_raster()

kf.env.trees <- envelope(trees_ppp, Kest, correction="border")
plot(kf.env.trees)

trees_LE20 <- trees[trees$dia<=20,]
trees_LE20_ppp <- ppp(trees_LE20$x, trees_LE20$y, c(0,200), c(0,200))
kf.env.treesLE20 <- envelope(trees_LE20_ppp, Kest, correction="border")
plot(kf.env.treesLE20)
```